Seismic parameter estimation and the Canadian crust

### 2013-02-15 paper/draftBP.tex
paper

%% ------------------------------------------------------------------------ %%
\documentclass[draft]{article}

@@

%% -----------------------------------------------%%
\section{Data and method}
   Seismic data utilized in this study comes from 343 stations across Canada from all available networks. Of these stations 146 were used for analysis. All stations are broadband with data being used between the years 2000 and 2012. Receiver function analysis has been widely employed over many years to investigate the earths structure.
   Two derivations have been used in this study, one well tested stacking approach (Zhu and Kanamori, 2000) and the other a recent method that has not yet been employed at scale (Bostock, 2010). Both methods rely on the fact that the incoming S-wave contains energy from the direct arrival as well as reflected phases resulting from sharp velocity contrasts. Deconvolution of the P-wave energy arrival, used as an estimate for the source function, from the S-wave signal produces a Green's function with energy peaks at the times of the main and reflected arrivals. For better separation of P and S wave energy all components are rotated into radial and transverse dimensions and transformed into P, SV and SH components with a wave field decomposition transfer matrix (Bostock, 1998).
   Deconvolution is performed by minimizing the general cross validation function $GCV(\delta)$ where $\delta$ is used the regularizer (Bostock, 1998; Golub et al., 1979). This method is an $L_2$ frequency domain deconvolution which performs quickly and makes no assumptions on the noise in the data. All deconvolved signals are filtered between 0.04hz and 3.0hz and are stacked with weights $w_1=0.5$, $w_1=0.3$, $w_1=0.2$ for the Ps, PpPs and PpSs phases respectively. Error calculations are performed with bootstrap resampling by rerunning the processing with randomly chosen receiver functions - allowing multiples - 1024 times. The error is obtained by taking the standard deviation of the results.

\subsection{Vp methods}
   The stacking approach outlined above provides estimates for the parameters $R=\frac{V_P}{V_S}$ as well as crustal thickness $H$ - which gives the depth of the Moho. A method which estimates $V_P$ and $V_S$ separately (Bostock, 2010) is also employed for select stations. This method makes use of the fact the dependence on $H$ in the travel time equations can be removed if we divide the reflected phases by the main arrival.
$$ t_{Pps}(p_i) = \frac{ \sqrt{R^2-p_i^2V_P^2} + \sqrt{1-p_i^2V_P^2} } {\sqrt{R^2-p_i^2V_P^2} + \sqrt{1-p_i^2V_P^2} } t_{Ps}(p_i) $$
$$ t_{Pss}(p_i) = \frac{2\sqrt{R^2-p_i^2V_P^2}} {\sqrt{R^2-p_i^2V_P^2} + \sqrt{1-p_i^2V_P^2} } t_{Ps}(p_i) $$
This has the advantage that no assumptions on $V_P$ are necessary to perform the gridsearch and stack. A similar stacking method as the Zhu and Kanamori approach may be employed for $t_{Pps}$ and $t_{Pss}$ as long as an estimate for $t_{Ps}$ exists. With estimates for $R$ and $V_P$ a simple line search along $H$ can be made using all three undivided travel time equations.

@@

@@

>>>>

%% ------------------------------------------------------------------------ %%
\documentclass[draft, 12pt]{article}
\usepackage[english]{babel}
\usepackage[final]{graphicx}
\graphicspath{ {figures/} }
\DeclareGraphicsExtensions{.png}

@@

%% -----------------------------------------------%%
\section{Data and method}
   Seismic data utilized in this study come from 343 stations across Canada from all available networks. Of these stations 146 are of sufficient quality to be useful for study. All stations are broadband with data selected between the years 2000 and 2012, for a total of more than 700 earthquake sources. All teleseismic events outside of the 30 to 100 degrees of epicentral distance are removed before processing. Data is selected for reasonable signal to noise ratioand sufficient impulsiveness that arrival times can be calculated. All data is analyzed using receiver function methods, an approach which has been widely employed over many years to investigate the earths structure.

   Two derivations of receiver function analysis have been used in this study, one well tested stacking approach and the other a recent method that has not yet been employed at scale. Both methods rely on the fact that the incoming S-wave contains energy from the direct arrival as well as reflected phases resulting from sharp velocity contrasts. Deconvolution of the P-wave energy arrival, used as an estimate for the source function, from the S-wave signal produces a Green's function with energy peaks at the times of the main and reflected arrivals. For better separation of P and S wave energy all components are rotated into radial and transverse dimensions and transformed into P, SV and SH components with a wave field decomposition transfer matrix (Bostock, 1998).
   Deconvolution is performed by minimizing the general cross validation function $GCV(\delta)$ where $\delta$ is used the regularizer (Bostock, 1998; Golub et al., 1979). This method is an $L_2$ frequency domain deconvolution which performs quickly and makes no assumptions on the noise in the data. All deconvolved signals are filtered between 0.04hz and 3.0hz and are stacked with weights $w_1=0.5$, $w_1=0.3$, $w_1=0.2$ for the Ps, PpPs and PpSs phases respectively. Error calculations are performed with bootstrap resampling by rerunning the processing with randomly chosen receiver functions - allowing multiples - 1024 times. The error is obtained by taking the standard deviation of the results.

\subsection{Vp/Vs method}
   A well tested approach for calculating $R=\frac{V_P}{V_S}$ and $H$ was employed by Zhu and Kanamori (2000). After deconvolution the signals are stacked and a gridsearch is performed using the travel time equations for different values of $R$ and $H$ and a spectrum of slowness. For each parameter combination the travel time functions output estimates for the differential arrival time of the S-component main and reflected phases. These times are used to pick the energy in each receiver function - with the large amplitudes at arrivals stacking to provide an estimate for the optimal parameter estimation.

\subsection{Vp method}
   The stacking approach outlined above provides estimates for the parameters $R=\frac{V_P}{V_S}$ as well as crustal thickness $H$ - which gives the depth of the Moho. A method which estimates $V_P$ and $V_S$ separately (Bostock, 2010) is also employed for select stations. This method makes use of the fact the dependence on $H$ in the travel time equations can be removed if we divide the reflected phases by the main arrival.
$$ t_{Pps}(p_i) = \frac{ \sqrt{R^2-p_i^2V_P^2} + \sqrt{1-p_i^2V_P^2} } {\sqrt{R^2-p_i^2V_P^2} + \sqrt{1-p_i^2V_P^2} } t_{Ps}(p_i) $$
$$ t_{Pss}(p_i) = \frac{2\sqrt{R^2-p_i^2V_P^2}} {\sqrt{R^2-p_i^2V_P^2} + \sqrt{1-p_i^2V_P^2} } t_{Ps}(p_i) $$
This has the advantage that no assumptions on $V_P$ are necessary to perform the gridsearch and stack. A similar stacking method as the Zhu and Kanamori approach may be employed for $t_{Pps}$ and $t_{Pss}$ as long as an estimate for $t_{Ps}$ exists. With estimates for $R$ and $V_P$ a simple line search along $H$ can be made using all three undivided travel time equations.

@@

\subsection{Vp database}
   Accompanying the processed estimates outlines above are data from controlled source experiments collected and compiled by external sources. The data was compiled by Mooney (personal communication, 2012).

@@

\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{thompsonComparison}
  \caption{A picture of a gull.}
\end{figure}
